### 2013-09-26 evaluation/paper.tex
smallcaps for datasets, some more corrections by vlad

We refer to the learning on one fold as OCR-small, and learning on nine folds as OCR-big.

@@

\subsection{Multi-Label Classification}
Multi-label classification is a generalization of multi-class classification,
in which each example can be associated with more than one class. In other
words, the algorithm must decide for each sample and for each class whether the
sample belongs to that class or not.
Multi-label classification was first formulated as a structured prediction problem
in by \citet{finley2008training}, who used to investigate the influence of
approximate inference on the $n$-slack cutting plane algorithm. In their formulation
each class is represented as a binary node in a factor graph---the states representing
presence or absence of the class. A different factor of pairwise potentials is introduced
between each pair of classes.
We also consider a different model, where pairwise potentials are only
introduced between specific nodes. We build a tree over the binary variables by
computing the Chow-Liu Tree~\citep{chow1968approximating} of the targets. While
this results in a less expressive model, a tree-shaped model allows for exact
inference via message passing.
We use two of the datasets used in \citet{finley2008training}, the scene
and yeast datasets. We choose these two as these are real-world datasets for
which the pairwise approach outlined above actually improves upon the
baseline~\citep{finley2008training}.
The scene dataset has six labels, 294 input features, 1211 training samples and
1196 test samples. This leads to $204 \cdot 6 = 1224$ parameters for the unary potentials,
$3 \cdot 5 \cdot 4 = 60$ parameters for the full pairwise potentials (four for each edge), and
$5 \cdot 4 = 20$ parameters for the pairwise potentials of the tree-shaped model.
Using four parameters for each edge is a slight over-parametrization that
simplifies writing out the model.
The yeast dataset has 14 labels, 103 features, 1500 training samples and 971
test samples. The resulting numbers of parameters are $14 \cdot 103=1442$ parameters
for the unary potentials, $7 \cdot 13 \cdot 4=364$ parameters for the full pairwise
potential, and $13 \cdot 4=52$ parameters for the tree-shaped model.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XXXXXXX@{}}
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0007}\\
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0007}
    \end{tabu}
\caption{%
    Visualization of the snakes dataset. The top row shows input patterns, the bottom row the corresponding
    labels. The colors in the image showing the labels correspond to dark blue for background, and encoding
    the length of the snake from head (red) to tail (blue).
\figlabel{snakes_illustration}}
\end{figure}

\subsection{2D Grid CRF (Snakes)}
The snakes dataset is a synthetic dataset where samples are labeled 2D grids.
It was introduced in \citet{nowozin2011decision} to demonstrate the importance
of learning conditional pairwise potentials. The dataset consists of ``snakes''
of length ten traversing the 2D grid. Each grid cell that was visited is marked
as the snake heading out towards the top, bottom, left, or right, while
unvisited cells are marked as background.  The goal is to predict a labeling of
the snake from ``head'' to ``tail'', that is, assigning numbers from zero to
nine to the cells that are occupied by the snake.  \Figref{snakes_illustration}
illustrates the principle.
Local evidence for the target label is weak except
around head and tail, making this a challenging task, requiring strong pairwise
potentials. The dataset is noise-free in the sense that given the above
description, a human could easily produce the desired labeling without making
any mistake. The dataset is also interesting as the model of
\citet{nowozin2011decision} produced notoriously hard-to-optimize energy
functions.

In \citet{nowozin2011decision}, the input was encoded into five RGB colors
(``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
more suitably for our linear methods, we convert this representation to a
one-hot encoding of the five states.

@@

\subsection{Superpixel CRFs for Semantic Segmentation}
Our main attention is devoted to the use of conditional random fields for
semantic segmentation.  We use the Pascal VOC and MSRC datasets.
The Pascal VOC has 964 training images, each divided into
around 100 superpixels. There are 20 object categories, and an additional background class. The % TODO check! refer to other part
MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
each. There are 21 classes in the MSRC-21 dataset.
Each superpixel is represented as an output variable, with the ground truth
obtained by majority vote over the pixels belonging to the superpixel. We
removed all superpixels in which the majority of pixels is labeled ``void'', leading to some samples
having much fewer than 100 variables.
Pairwise potentials are introduced for each pair of neighboring superpixels.
We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC-21 dataset.
We use the same three pairwise features for both datasets: a constant feature, a color
feature and a feature encoding relative vertical position.
Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
2646$ parameters for the potentials on the MSRC-21 dataset.

@@

\begin{table}
    \begin{tabu} to \linewidth{@{}X[1,l]YX[1,c]X[1,c]YY@{}}
    %\multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\
    \toprule
    Dataset    & {samples} & variables & graph & {dim $\theta$} & {labels}\\
    \cmidrule{1-6}
    MNIST      & 60000   & 1     & none  & 7840  & 10\\
    OCR-small  & 704     & 3--14 & chain & 1444  & 26\\
    OCR-large  & 6173    & 3--14 & chain & 1444  & 26\\
    scene-tree & 1211    & 6     & tree  & 1244  & 2\\
    scene-full & 1211    & 6     & loopy & 1284  & 2\\
    yeast-tree & 1500    & 14    & tree  & 1494  & 2\\
    yeast-full & 1500    & 14    & loopy & 1806  & 2\\
    snake      & 200     &84--168& loopy & 22275 & 11\\
    MSRC-21     & 276     & 7--113& loopy &  1764 & 21\\
    Pascal VOC & 964     & 8--112& loopy &  2646 & 21\\
    \bottomrule
    \end{tabu}
    \caption{Summary of datasets used in the evaluation. \tablabel{datasets}}
\end{table}

@@

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these tasks, exact inference is
possible, and we can compute the exact objective of the various algorithms
easily. We use the dynamic programming algorithm implemented in \textsc{OpenGM}
for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with
the AD$^3$ algorithm. %FIXME AD3 solver description?
We found that using the linear programming relaxation on the scene dataset is
often tight, and there was no need to resort to approximate inference. Inference
on the yeast dataset was more complex, and on the verge of being non-practical.
We use this dataset as an example for very costly inference.

@@

We now consider datasets with non-trivial inference.
The results for OCR-small are shown in \Figref{ocr_small_results}.
Considering the plot against passes over the dataset on the left, several
trends can be observed: the $n$-slack cutting plane algorithms converge
fastest, with the version that recomputes the QP at every step leading the
race. The algorithms are closely followed by the caching $1$-slack cutting
plane algorithm.  They are followed by the significantly slower SZLJSP
algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
subgradient descent.  These results are intuitive, as they reflect ``how much
work'' each algorithm does for each loss-augmented prediction step. More work
towards the objective leads to faster convergence.  This ``more work'' is
quantified on the right hand side of \Figref{ocr_small_results}, where the suboptimality is plotted against time.
Clearly the $n$-slack algorithms do ``too much'' work, leading to very slow
convergence.  Also, caching does not speed up learning learning on this
dataset.
%
As on MNIST, the $1$-slack algorithm stops before the SZLJSP (in terms of
iterations) because of a tighter dual.  A detailed view is shown in
\Figref{ocr_small_dual} on the right.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/mnist_duality_gap}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_duality_gap}
    \end{tabu}
\caption{%
    Primal objective and dual objective (dashed lines) for SZLJSP and $1$-slack cutting plane on MNIST (left) and OCR-small (right).
\figlabel{ocr_small_dual}}
\end{figure}

@@

The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
algorithm when solving the QP at every step here.
The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
with respect to runtime. The SZLJSP algorithm converges fast initially, but the $1$-slack cutting plane
is faster at high precision and faster in certifying the desired duality gap.
\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full_time}
    \end{tabu}
\caption{%
   Convergence of the primal suboptimality on the multi-label datasets. From top to
   bottom: the scene dataset using a tree model, the scene dataset using a full
   model, the yeast dataset using a tree model, and the yeast dataset using a
   full model.
\figlabel{multi-label-results}}
\end{figure}

Next, we consider the multi-label task.  The results for the yeast and scene
datasets are shown in \Figref{multi-label-results}. The caching $1$-slack
algorithm is much faster than the non-caching one on both datasets and both
graph structures.
While the two variants of the $n$-slack algorithm are equally fast with respect
to passes over the training data, solving the QP every 100 steps is among the
fastest algorithms, while solving the QP at every step is among the slowest.
We applied this method only to the smaller scene dataset, as applying it on the
yeast dataset was not practicable.

@@

\subsection{Experiments using Approximate Inference}
For the remaining tasks, the snakes dataset and image segmentation, exact
inference is intractable for learning.
%
We use two different approximate inference algorithms, fusion moves, a fast
local search procedure, and the linear programming relaxation provided by AD$^3$
not using branch-and-bound, in contrast to the multi-label setup above.

@@

\Figref{snakes_results} shows the results on the snakes dataset.
The top row shows the objective when learning with fusion moves, the next row
the training set loss.  First, we observe that using fusion moves, even the
approximate objective does not reach the desired optimality gap for the
$1$-slack and SZLJSP algorithms. This is caused by the algorithms not finding
any further constraints. The $n$-slack algorithm fares a bit better and
achieves a higher dual value. Looking at the training set loss, it is clear that the stochastic
algorithms have an advantage, possibly caused by not stopping too early.
Still, all algorithms ultimately fail to solve the task.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}
    \end{tabu}
\caption{%
    Results on the snakes dataset. The top four plots show results using fusion
    move inference, the bottom four plots using AD$^3$. Dashed lines indicate the dual objective.
\figlabel{snakes_results}}
\end{figure}

The last set of experiments is on the segmentation dataset MSRC-21 and Pascal
VOC\@.  Because of the large number of labels and samples, it is only feasible
to use the fusion move inference algorithm. The results are shown in
\Figref{msrc_pascal_curves}, with the top four plots showing results on
MSRC-21, and the bottom four plots showing results on Pascal VOC\@.
First, we notice that all algorithms achieve low (approximate) duality gaps, and none
of the algorithms terminates prematurely. Also, all algorithms achieve a
similar dual objectives and similar training errors after convergence.
In both datasets, the $1$-slack algorithm is somewhat slow in minimizing the
training set error, and the $n$-slack and stochastic algorithms are much
faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
down learning on both datasets. This is in contrast to the yeast datasets, where the converse was true,
even though yeast has a similar number of training examples to Pascal VOC.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time_loss}
    \end{tabu}
\caption{%
    Objective and traning set loss on the segmentation task. The top four plots
    show results for the MSRC-21 dataset, the bottom four plots for the Pascal
    VOC dataset.
\figlabel{msrc_pascal_curves}}
\end{figure}

@@

With respect to approximate learning, we found that the inference algorithm has
a stronger impact on the performance than the learning algorithm.
While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes dataset than
the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
task nearly perfectly.
For the segmentation datasets, it is harder to judge the outcome, though there
seems to be little difference between the learning algorithms in terms of the
final result.  In the next chapter, we will demonstrate how we can learn
exactly, even with the complex models used for semantic segmentation.

>>>>

We refer to the learning on one fold as \textsc{OCR-small}, and learning on nine folds as OCR-big.

@@

\subsection{Multi-Label Classification}
Multi-label classification is a generalization of multi-class classification,
in which each example can be associated with more than one class. In other
words, the algorithm must decide for each sample and for each class whether the
sample belongs to that class or not.
Multi-label classification was first formulated as a structured prediction problem
by \citet{finley2008training}, who used to investigate the influence of
approximate inference on the $n$-slack cutting plane algorithm. In their formulation
each class is represented as a binary node in a factor graph---the states representing
presence or absence of the class. A different factor of pairwise potentials is introduced
between each pair of classes.
We also consider a different model, where pairwise potentials are only
introduced between specific nodes. We build a tree over the binary variables by
computing the Chow-Liu Tree~\citep{chow1968approximating} of the targets. While
this results in a less expressive model, a tree-shaped model allows for exact
inference via message passing.
We use two of the datasets used in \citet{finley2008training}, the \textsc{scene}
and \textsc{yeast} datasets. We choose these two as these are real-world datasets for
which the pairwise approach outlined above actually improves upon the
baseline~\citep{finley2008training}.
The \textsc{scene} dataset has six labels, 294 input features, 1211 training samples and
1196 test samples. This leads to $204 \cdot 6 = 1224$ parameters for the unary potentials,
$3 \cdot 5 \cdot 4 = 60$ parameters for the full pairwise potentials (four for each edge), and
$5 \cdot 4 = 20$ parameters for the pairwise potentials of the tree-shaped model.
Using four parameters for each edge is a slight over-parametrization that
simplifies writing out the model.
The \textsc{yeast} dataset has 14 labels, 103 features, 1500 training samples and 971
test samples. The resulting numbers of parameters are $14 \cdot 103=1442$ parameters
for the unary potentials, $7 \cdot 13 \cdot 4=364$ parameters for the full pairwise
potential, and $13 \cdot 4=52$ parameters for the tree-shaped model.

\begin{figure}
    \begin{tabu} to \linewidth{@{}XXXXXXX@{}}
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0007}\\
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0001}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0002}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0003}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0004}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0005}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0006}&%
    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0007}
    \end{tabu}
\caption{%
    Visualization of the \textsc{snakes} dataset. The top row shows input patterns, the bottom row the corresponding
    labels. The colors in the image showing the labels correspond to dark blue for background, and encoding
    the length of the snake from head (red) to tail (blue).
\figlabel{snakes_illustration}}
\end{figure}

\subsection{2D Grid CRF (Snakes)}
The \textsc{snakes} dataset is a synthetic dataset where samples are labeled 2D grids.
It was introduced by \citet{nowozin2011decision} to demonstrate the importance
of learning conditional pairwise potentials. The dataset consists of ``snakes''
of length ten traversing the 2D grid. Each grid cell that was visited is marked
as the snake heading out towards the top, bottom, left, or right, while
unvisited cells are marked as background.  The goal is to predict a labeling of
the snake from ``head'' to ``tail'', that is, assigning numbers from zero to
nine to the cells that are occupied by the snake.  \Figref{snakes_illustration}
illustrates the principle.
Local evidence for the target label is weak except
around head and tail, making this a challenging task, requiring strong pairwise
potentials. The dataset is noise-free in the sense that given the above
description, a human could easily produce the desired labeling without making
any mistake. The dataset is also interesting as the model proposed by
\citet{nowozin2011decision} produced notoriously hard-to-optimize energy
functions.

Originally, the input is encoded into five RGB colors
(``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
more suitably for our linear methods, we convert this representation to a
one-hot encoding of the five states.

@@

\subsection{Superpixel CRFs for Semantic Segmentation}
Our main attention is devoted to the use of conditional random fields for
semantic segmentation.  We use the \textsc{Pascal VOC} and MSRC datasets.
The \textsc{Pascal VOC} has 964 training images, each divided into
around 100 superpixels. There are 20 object categories, and an additional background class. The % TODO check! refer to other part
MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
each. There are 21 classes in the MSRC-21 dataset.
Each superpixel is represented as an output variable, with the ground truth
obtained by majority vote over the pixels belonging to the superpixel. We
removed all superpixels in which the majority of pixels is labeled ``void'', leading to some samples
having much fewer than 100 variables.
Pairwise potentials are introduced for each pair of neighboring superpixels.
We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
21 unary features for the \textsc{Pascal VOC} dataset and 63 unary features for the MSRC-21 dataset.
We use the same three pairwise features for both datasets: a constant feature, a color
feature and a feature encoding relative vertical position.
Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
for the potentials on \textsc{Pascal VOC} and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
2646$ parameters for the potentials on the MSRC-21 dataset.

@@

\begin{table}
    \begin{tabu} to \linewidth{@{}X[1,l]YX[1,c]X[1,c]YY@{}}
    %\multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\
    \toprule
    Dataset             & {samples} & variables & graph & {dim $\theta$} & {labels}\\
    \cmidrule{1-6}
    MNIST               & 60000   & 1     & none  & 7840  & 10\\
    \textsc{OCR-small}  & 704     & 3--14 & chain & 1444  & 26\\
    \textsc{OCR-large}  & 6173    & 3--14 & chain & 1444  & 26\\
    \textsc{scene-tree} & 1211    & 6     & tree  & 1244  & 2\\
    \textsc{scene-full} & 1211    & 6     & loopy & 1284  & 2\\
    \textsc{yeast-tree} & 1500    & 14    & tree  & 1494  & 2\\
    \textsc{yeast-full} & 1500    & 14    & loopy & 1806  & 2\\
    \textsc{snakes}     & 200     &84--168& loopy & 22275 & 11\\
    MSRC-21             & 276     & 7--113& loopy &  1764 & 21\\
    \textsc{Pascal VOC} & 964     & 8--112& loopy &  2646 & 21\\
    \bottomrule
    \end{tabu}
    \caption{Summary of datasets used in the evaluation. \tablabel{datasets}}
\end{table}

@@

\subsection{Experiments using Exact Inference}
In this section, we present results on multi-class classification, sequence
labeling and multi-label prediction. In these tasks, exact inference is
possible, and we can compute the exact objective of the various algorithms
easily. We use the dynamic programming algorithm implemented in \textsc{OpenGM}
for experiments for the tree and chain models.
For the full pairwise multi-label model, we use branch-and-bound together with
the AD$^3$ algorithm. %FIXME AD3 solver description?
We found that using the linear programming relaxation on the \textsc{scene} dataset is
often tight, and there was no need to resort to approximate inference. Inference
on the \textsc{yeast} dataset was more complex, and on the verge of being non-practical.
We use this dataset as an example for very costly inference.

@@

We now consider datasets with non-trivial inference.
The results for \textsc{OCR-small} are shown in \Figref{ocr_small_results}.
Considering the plot against passes over the dataset on the left, several
trends can be observed: the $n$-slack cutting plane algorithms converge
fastest, with the version that recomputes the QP at every step leading the
race. The algorithms are closely followed by the caching $1$-slack cutting
plane algorithm.  They are followed by the significantly slower SZLJSP
algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
subgradient descent.  These results are intuitive, as they reflect ``how much
work'' each algorithm does for each loss-augmented prediction step. More work
towards the objective leads to faster convergence.  This ``more work'' is
quantified on the right hand side of \Figref{ocr_small_results}, where the suboptimality is plotted against time.
Clearly the $n$-slack algorithms do ``too much'' work, leading to very slow
convergence.  Also, caching does not speed up learning learning on this
dataset.
%
As on MNIST, the $1$-slack algorithm stops before the SZLJSP (in terms of
iterations) because of a tighter dual.  A detailed view is shown in
\Figref{ocr_small_dual} on the right.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/mnist_duality_gap}&%
    \includegraphics[width=\linewidth]{evaluation/images/letters_small_duality_gap}
    \end{tabu}
\caption{%
    Primal objective and dual objective (dashed lines) for SZLJSP and $1$-slack cutting plane on MNIST (left) and \textsc{OCR-small} (right).
\figlabel{ocr_small_dual}}
\end{figure}

@@

The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
algorithm when solving the QP at every step here.
The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
with respect to runtime. The SZLJSP algorithm converges fast initially, but the $1$-slack cutting plane
is faster at high precision and faster in certifying the desired duality gap.
\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full}&%
    \includegraphics[width=\linewidth]{evaluation/images/yeast_full_time}
    \end{tabu}
\caption{%
   Convergence of the primal suboptimality on the multi-label datasets. From top to
   bottom: the \textsc{scene} dataset using a tree model, the \textsc{scene} dataset using a full
   model, the \textsc{yeast} dataset using a tree model, and the \textsc{yeast} dataset using a
   full model.
\figlabel{multi-label-results}}
\end{figure}

Next, we consider the multi-label task.  The results for the \textsc{yeast} and \textsc{scene}
datasets are shown in \Figref{multi-label-results}. The caching $1$-slack
algorithm is much faster than the non-caching one on both datasets and both
graph structures.
While the two variants of the $n$-slack algorithm are equally fast with respect
to passes over the training data, solving the QP every 100 steps is among the
fastest algorithms, while solving the QP at every step is among the slowest.
We applied this method only to the smaller \textsc{scene} dataset, as applying it on the
\textsc{yeast} dataset was not practicable.

@@

\subsection{Experiments using Approximate Inference}
For the remaining tasks, the \textsc{snakes} dataset and image segmentation, exact
inference is intractable for learning.
%
We use two different approximate inference algorithms, fusion moves, a fast
local search procedure, and the linear programming relaxation provided by AD$^3$
not using branch-and-bound, in contrast to the multi-label setup above.

@@

\Figref{snakes_results} shows the results on the \textsc{snakes} dataset.
The top row shows the objective when learning with fusion moves, the next row
the training set loss.  First, we observe that using fusion moves, even the
approximate objective does not reach the desired optimality gap for the
$1$-slack and SZLJSP algorithms. This is caused by the algorithms not finding
any further constraints. The $n$-slack algorithm fares a bit better and
achieves a higher dual value. Looking at the training set loss, it is clear that the stochastic
algorithms have an advantage, possibly caused by not stopping too early.
Still, all algorithms ultimately fail to solve the task.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}
    \end{tabu}
\caption{%
    Results on the \textsc{snakes} dataset. The top four plots show results using fusion
    move inference, the bottom four plots using AD$^3$. Dashed lines indicate the dual objective.
\figlabel{snakes_results}}
\end{figure}

The last set of experiments is on the segmentation dataset MSRC-21 and \textsc{Pascal
VOC}.  Because of the large number of labels and samples, it is only feasible
to use the fusion move inference algorithm. The results are shown in
\Figref{msrc_pascal_curves}, with the top four plots showing results on
MSRC-21, and the bottom four plots showing results on \textsc{Pascal VOC}.
First, we notice that all algorithms achieve low (approximate) duality gaps, and none
of the algorithms terminates prematurely. Also, all algorithms achieve a
similar dual objectives and similar training errors after convergence.
In both datasets, the $1$-slack algorithm is somewhat slow in minimizing the
training set error, and the $n$-slack and stochastic algorithms are much
faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
down learning on both datasets. This is in contrast to the \textsc{yeast} datasets, where the converse was true,
even though \textsc{yeast} has a similar number of training examples to \textsc{Pascal VOC}.

@@

\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time}\\
    \includegraphics[width=\linewidth]{evaluation/images/pascal_loss}&%
    \includegraphics[width=\linewidth]{evaluation/images/pascal_time_loss}
    \end{tabu}
\caption{%
    Objective and traning set loss on the segmentation task. The top four plots
    show results for the MSRC-21 dataset, the bottom four plots for the \textsc{Pascal
    VOC} dataset.
\figlabel{msrc_pascal_curves}}
\end{figure}

@@

With respect to approximate learning, we found that the inference algorithm has
a stronger impact on the performance than the learning algorithm.
While the SSGD and SZLJSP algorithms have somewhat lower training error on the \textsc{snakes} dataset than
the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
task nearly perfectly.
For the segmentation datasets, it is harder to judge the outcome, though there
seems to be little difference between the learning algorithms in terms of the
final result.  In the next chapter, we will demonstrate how we can learn
exactly, even with the complex models used for semantic segmentation.
